---
title: "441 Homework 1"
author: ''
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  pdf_document:
    toc: yes
---

## Question 1 (20 points)
### a
What is the sample regression equation relating NUMPORTS to COST?

```{r 1a}
Q1 <- read.csv("COMMNODES.csv")

#### Correlation Coefficient and Scatterplot ####
plot(Q1$NUMPORTS, Q1$COST, xlab = "Numports", ylab = "Cost")
abline(lm(Q1$COST ~ Q1$NUMPORTS))

cor(Q1$COST, Q1$NUMPORTS)

#### Regression Model ####
lm.COMM <- lm(Q1$COST ~ Q1$NUMPORTS)
anova(lm.COMM)
sum <- summary(lm.COMM)
sum$coefficients
beta_0 <- sum$coefficients[1]
beta_1 <- sum$coefficients[2]
beta_0
beta_1
```

The regression equation is Cost = 16593.65 + 650.1692*Numports

### b 
Is there sufficient evidence to conclude that a linear relationship exists between COST and NUMPORTS?

Test to use: person

H0:ρ=0(No correlation)

H1:ρ≠0(There is a correlation)
```{r 1b}
cor.test(Q1$NUMPORTS, Q1$COST,alternative = "two.sided", method = "pearson")
```
INTERPRETATION AND FINAL CONCLUSION: At 5% significance level, there is sufficient evidence to support the claim that a linear relationship exists between COST and NUMPORTS.

### c 
Find and interpret the 95% confidence interval estimate of β1.
```{r 1c}
confint(lm.COMM)
```
The 95% confidence interval estimate of β1 is (504.3763, 795.962). We are 95% confident that the interval between 504.4 dollars and 796.0 dollars contains the true population mean change in cost of each additional port.

### d
A claim has been made that each new access port adds at least $1,000 to the installation cost of a communications node. Test this claim by using the test of β1.

This requires us to test whether beta1 < 1000

H0:beata1 >= 1000

H0:beata1 < 1000
```{r 1d}
####  t_critical value
t_critical <- qt(0.05, nrow(Q1)-1-1, lower.tail = T)
t_critical

#### t = (b - beta)/std. error
std <- sum$coefficients[4]
t_test_stat <- (beta_1 - 1000)/std
t_test_stat

t_test_stat < t_critical
```
INTERPRETATION AND FINAL CONCLUSION: At 5% significance level, since the t_test_stat is less than the t_critical, for a left-tailed test we reject H0: beta1 >= 1000. We infer at 5% significance level that beta1 < 1000 is statistically significant, which means the claim that each new access port adds at least $1,000 to the installation cost of a communications node is not true.

## Question 2 (20 points)

### a
Run a multiple regression analysis and discuss your findings.
```{r 2a}
Q2 <- read.csv("MedicalCo.csv")
SALES <- Q2$SALES
ADV <- Q2$ADV
BONUS <- Q2$BONUS
MKTSHR <- Q2$MKTSHR
COMPET <- Q2$COMPET

# Correlation Matrix
cor(as.matrix(Q2[1:3]))

# Multiple Regression Model
Model1 <- lm(SALES ~ ADV + BONUS)
anova(Model1)
summary(Model1)



```
As we can see from the correlation matrix, there is a strong positive correlation between sales and ADV, 0.9.
Also a positive correlation between sales and bonus, 0.568.

The regression eaquation is : SALES = -516.4443 + 2.4732* ADV + 1.8562*BONUS
beta0 =-516.4. It means that when there is no advertising and bonus, sales would be -516.4 thousand dollars.This is a meaningless intercept because sales will never be negative.

beta1 = 2.4732. It means that for every 100 dollars increase of ADV, sales would increase 2473.2 dollars.

beta2 = 1.8562. It means that for every 100 dollars increase of bonus, sales would increase 1856.2 dollars.
The R-square value is 85.49% which means 85.49% of variation in SALES is explained by the model. The remaining 14.51% is unexplained. The adjusted R-square is 84.18%. The standard error of the model is 90.74 thousand dollars.



### b
Run an appropriate test to see if these two variables should be included in the model or not. What is your conclusion?
```{r 2b}
# Correlation Matrix
cor(as.matrix(Q2[c(1,4,5)]))

# Multiple Regression Model
Model2 <- lm(SALES ~ ADV + BONUS + MKTSHR + COMPET)
anova(Model2)
summary(Model2)

#comparison
anova(Model1, Model2)
```
As we can see from the correlation matrix, there is a weak positive correlation between sales and MKTSHR, 0.0231.
Also a weak positive correlation between sales and COMPET, 0.377.

From the summary of this model, the p-value of MKTSHR and COMPET are pretty large, which means these two variables are not significant. Also, p-value of partial F is more than 0.05.
It's better to exclude these two variables and choose model1.





